\thetitle

### 2021-02-10 writing/reviews/pldi-rebuttal-high-level-first.md

> Theoretical gradual type systems usually allow programmers to control
> the precision of type annotations in a fine-grained way with the Dyn
> type rather than distinguishing typed and untyped modules. Would the
> same evaluation method be able to directly appliable to blame in such
> systems?

Yes. Incorporating type Dyn boils down to considering a hierarchy of types
for each component rather than a single type. At the same time, our method
  is parametric to the "size" of a component and adapts in a
  straightforward manner to a setting where each definition can be
  considered a component instead of a whole module. Of course both of
  these adjustments lead to significantly larger scenario lattices than
  the ones we investigate, so keeping the experiment computationally
  feasible would require more aggressive sampling.

> Please explain what impedance mismatches are in the paper. Is it simply
> type mismatches?

"Impedence mismatch" is our name for a run-time disagreement between a
type annotation and an untyped value. The point of the name is to suggest
that either side (type or value) could be at fault. We will adjust the prose.

@@

You are right that for a given trail of length $n$ the probability that
the random rational programmer discovers the buggy component on every step
is $1/n$.  The missing piece is that we randomly sample from the set of
scenarios -- and by extension the set of trails. So the shape of the
random programmer's effort matches our sampling distribution. That said,
as you observe, the distribution of the random programmer's effort is not
relevant to our results, so we will remove the characterization.

> * page 9 line 980: Figure 5.4 -> Figure 6
> * page 10 Figure 7: Null -> Random
> * page 11 line 1103: ( 770 -> (770

Thank you. We will fix all these.

Review #66C
===========================================================================

> Paper summary
> -------------
>
> [....]
>
> To evaluate this approach, the paper also presents empirical results. Starting
> with a small corpus of programs, the authors use mutations to systematically
> (but synthetically) introduce type errors -- e.g., one mutation replaces a
> constants with another of a different type, and another mutation swaps method
> identifiers. They then conduct experiments to quantify the number of steps
> required by a rational programmer for several blame assignment schemes: Natural
> (used in Typed Racket, which tracks at most one location for blame), Transient
> (used in Reticulated Python, which tracks a set of locations), and Erasure
> (which does not actually assign blame, and is widely used in industry).

chrdimo: TODO
effort is part 2/2 of the metrics

> But while there is a lot to like about the paper, it also has some
> shortcomings. The suite of programs used for the evaluation is fairly
> small (c.f., the SIGPLAN checklist).

The paper does follow the SIGPLAN checklist guidelines for Principled Benchmark
Choice:

- This paper is the first to conduct an empirical study for blame and gradual
  typing, so there is no established benchmark suite to reuse.

- The paper thus explains how we adapted a suite of gradual typing performance
  benchmarks to a new setting (section 4).

- The paper also explains how we selected pertinent benchmarks to work within
  our resource constraints.

We cannot find any minimum size guidelines in the SIGPLAN checklist.

@@

> The faults are introduced using synthetic mutations that may or may not
> correspond to the kinds of errors that arise in practice. This can be
> seen from the results in Figure 7 -- the lengths of the "trails" is
> quite small.

There is no existing catalog of bugs in gradually typed programs, nor any
statistics about common trail lengths.  Furthermore, most type-level bugs do
not survive  decent testing and thus do not make it to code repos. There are a
few anecdotes from the literature but they are mostly artificial. Hence, since
we need a large number of buggy scenarios for a meaningful study, a synthetic
approach is the only way forward. In addition, mutation allows us to tune the
set of scenarios to get a diverse set of interesting bugs.

@@

> While the study is mostly well done, it only considers two blame
> assignment strategies.

We would like to point out that the experiment considers six different
modes of the rational programmer spread across three different gradual typing
systems plus the random rational programmer. Figure 7 clearly shows them all.

@@

Review #66D
===========================================================================

> 2. That Transient --- as implemented in Typed Racket --- is much
> slower than Natural. ...
> Given that the experiments are developed and run in the context of the
> large, real-world Racket ecosystem, I wonder whether about the chance of
> incidental implementation choices or bugs affecting this result.

A side note here: Transient Typed Racket is faster than Natural Typed
Racket when we turn off blame. Exactly as a sanity check that this is not
an artifact of our implementation we have coded up benchmarks in Python
and we confirmed that in Reticulated the performance of Transient with
blame is even worse than what we have observed in Typed Racket.

> *Natural and Transient Under the Same Roof (Section 3)*
>
> This seems to be "only" an engineering challenge.
>
> [...]
>
> *Custom Mutators (Section 4)* The paper suggests that the mutators
> required here are significantly different and more subtle, but this does
> not seem to be the case when comparing Figure 3 and Table 3 of the POPL
> 2020 paper.  ...
>
> [...]
>
> *Debugging Strategies (Section 5)*
>
> These also seem to be overemphasized. The main design choice seems to be
> in how to use Transient blame, which reports multiple components, and
> picking the first and last are pretty intuitive.
>
> I'm also not sure why the exception mode is needed in addition to
> Erasure (I was confused in Section 2 L204-215 and Section 5.1 L732).
> Indeed, Section 5.3 defines the Erasure mode the follow the Natural
> exceptional mode.

@@

Thank you. We will fix these.

Review #66E
===========================================================================

> However, I don't find the results to be as conclusive as the authors
> state. Aside from the Erasure case, whose relevance I found unclear (see
> below), conclusions rest on some small-valued "more useful than"
> percentages. These are pretty hard to interpret, because the
> "usefulness" metrics necessarily build in a lot of simplifying
> assumptions. E.g. in 5.4, the "percentage of scenarios where ... is more
> effective" doesn't account for *extent* of difference, and the
> trail-length "programmer effort" metric in 5.5 is also unlikely to be a
> great proxy. These limitations are understandable, method-wise, but
> overall the results are fairly null... they let us continue believing
> "what we'd expect" but not with any great sense of added confidence.
> That doesn't entirely diminish the work, of course, but I'd say the
> current write-up overplays its hand a little. I appreciate the
> discussion of threats to validity.

TODO chrdimo
- see general comments for how to interpret results, their significance, and how they challenge existing perceptions
- usefulness asks if one succeeds when another failed; no extent for this
- if both succeed, can quantify difference, use trail length to do so
- trail length is proxy, paper discusses limitations, allows automation ... no humans + large scale

> The presentation of the results is very much around aggregates and summaries.
> Indeed the whole method is about having run a huge compute job over a large
> number of variants. Perhaps it's paranoia but I'd be interested to see some
> specific examples walked through by hand (in the paper) and some smaller sample
> manually classified (as results). That would add an extra sanity check that the
> metrics do correspond to some meaningful reality.

TODO chrdimo
....

if you are asking whether our results

@@

> line 313: it threw me that a program might not fail but "produce a wrong
> result". If it could be caught by the gradual type system, why can't it
> be caught at run time? My best get at explaining this is by what I wrote
> above, i.e. it's a consequence of Typed Racket's more refined static
> notion of type. In certain other systems this wouldn't be possible,
> because the static checker would only catch (albeit earlier) errors that
> would be caught at run time, so there would be no basis to call the
> result "wrong".

Consider a function with a type annotation that says its a function from
integers to string that formats its argument to a string and returns it.
If we import the function in an untyped module under Erasure there are no
checks to make sure that the function is going to be applied to integers.
So we can apply it to character or a string and still get back a result
except that it is not the string representation of a number.

> line 380: from this I inferred that "migratory typing" means "gradual
> typing applied at modulewise granularity". Assuming that's correct, it's
> worth saying directly.

We will clarify. Migratory typing is not about granularity, but whether the
type system forces rewrites upon untyped code. A gradual type system for Python
would be migratory. A new gradual-from-the-start language would not be.

@@

> line 424: "fully typed correct programs" -- presumably your method could
> also work with not-yet-fully-typed correct programs, just not ranging
> over the entire lattice in those cases. I was wondering whether that
> might give different/interesting results.

Yes, the approach could be applied.
ll: worth a sentence saying that this would probably not be worthwhile to do?
bg: TODO try explaining in 2-3 words

@@

> line 160: "disparity" is a strange word here. "diversity"?

> line 170: "Otherwise, results from..." -- the point was already clear

> line 174: ... instead of just saying what you built on, first say what
> you did! It's really not clear at this point.

> line 199: "blame set as another form of a stack trace" -- yes. I was hoping to
> see a clearer example of debugging with and without blame, i.e. something
> making explicit the similarities and differences between having blame info and
> having only a stack trace at the error site.

### 2021-02-10 writing/reviews/pldi-rebuttal-high-level-first.md

PLDI 2021 Paper #66 Author Response
===========================================================================

Thank you to all the reviewers for the helpful feedback.

The first part of our response addresses a couple of common themes in the reviews; the second part addresses all remaining concerns individually.

Are the empirical results conclusive or surprising?
===========================================================================

Here are the two relevant major conclusions from our experiment:

1) Transient and Natural make rational programmers more effective than
Erasure;

2) According to the rational programmer, the effectiveness of Natural is (merely)
comparable to Transient.

As review `D` astutely points out, the second one is surprising: _While Theory predicts that Natural blame is more accurate than Transient blame, the experimental results do not support this prediction to a sufficient degree._

Thus our results point to two possible explanations:

(i) a need for theoretical models with improved predictive power

The predictive power of gradual typing formal models fails when it comes to blame. In all likelihood, the current models are too small and do not include crucial
features. (It is posisble that the current models are inherently limited.)

(ii) a need for "blame data" from large systems

The presented data is based on systems with up to 15 modules, totalling 25,000 loc. While the chosen benchmarks represent a wide spectrum of realistic code, Natural may prove its worth only when the dependency chains are even deeper than the ones in the existing benchmarks.

In both cases, our new method provides the tools for validating new theoretical results and/or gathering new data.

Additionally, the experiment suggests that implementors can experiment with checks and blame to find a sweet spot between type soundness, blame effectiveness and performance. Our method
provides an automatic tool to guide their exploration.

Is the method novel? 
===========================================================================

### 2021-02-10 writing/reviews/pldi-rebuttal-high-level-first.md

The first part of our response addresses a couple of common themes in the reviews; the second part addresses all remaining concerns individually.

Are the empirical results conclusive or surprising?
===========================================================================

Here are the two relevant major conclusions from our experiment:

1. Transient and Natural make rational programmers more effective than
Erasure;

2. According to the rational programmer, the effectiveness of Natural is (merely)
comparable to Transient.

@@

As review `D` astutely points out, the second one is surprising: _While Theory predicts that Natural blame is more accurate than Transient blame, the experimental results do not support this prediction to a sufficient degree._

Thus our results point to two possible explanations:

- a need for theoretical models with improved predictive power

@@

In both cases, our new method provides the tools for validating new theoretical results and/or gathering new data.

Additionally, the experiment suggests that implementors can experiment with checks and blame to find a sweet spot between type soundness, blame effectiveness and performance. Our method
provides an automatic tool to guide their exploration.

Is the method novel? 
===========================================================================

Our inspiration is Lazarek et al.'s POPL paper, which shows the possibility of large-scale PL desiggn evaluations that do _not_ involve human sw devs.

The challenge is how to adapt this work to different problems than contracts. Given the close relationship between Natural and contracts, it is _natural_ to pursue an adaptation of the Lazarek method to gradual typing. Adaptation means retaining some ideas and injecting entirely new ones:

1. Lazarek et al.'s work can use off-the-shelf mutators because their programs are untyped. For typed prorgams, we must create new ones, for which we occasionally chose the same name as Lazarek --- a significant articulation mistake on our side. For the final version, we will use distinct names.

2. Lazarek et al.'s method cannot really conclude that blame is the reason for the success of the debugging process because it fails to isolate the effect of blame from that of checks or mere luck. What they can assert is only that blame shifting works. Our work demands a proper `control', which we have developed.

3. Lazarek et al. deal with a single contact system and examine a yes-or-no question, namely, whether  blame-shifting finds the bug for all debugging scenarios of a benchmark. Our experiment compares fundamentally different error-reporting schemes.

As to the specific complaints in reviews `c` and `d`:

* Natural and Transient under the same roof (section 3).

@@

Implementing Transient Typed Racket is not mere engineering.  Neither
Reticulated nor the Transient model provide answers about what shallow
checks correspond to Typed Racket's sophisticated types (e.g.
polymorphism, row types for mixins, occurrence typing) and what
information is necessary to record in the blame map. Moreover in the
process of the implementation we discovered issues with the published
Transient model and Reticulated that we had to fix. We do not think that
this paper is the appropriate place to describe the implementation in
detail.

* Custom mutators (section 4).

The identification of good mutators for gradual typing is not
straightforward. A bad mutator results in a type-level bug that is trivial
to locate. For instance, a bug that leads to a type-level error during the
evaluation of the component that contains it is trivial to locate. In
contrast, a bug that is non-trivial to locate is one that causes the buggy
component to evaluate without an error but produce instead a value of the
wrong type. Furthermore an interesting bug is one where the value with the
wrong type manages also to cross undetected to at least another component
before a type-level check flags it. Good mutators should be able to
produce these interesting bugs in our benchmarks. Finally, they should
produce interesting bugs that are detectable by all three systems so that
we can try all systems on the same debugging scenarios.  In general, there
is no methodology for constructing a new set of mutators that produce
interesting bugs for any definition of interesting. So narrowing down the
ones in the paper required rounds of trials to end up with a set of
mutators that leads to a sufficiently large set of interesting and diverse
bugs.

* TODO phrasing of "original set we tried" seems off
The sixteen mutators we include in the paper are not the original set we
tried and others that seemed like good candidates proved to be
ineffective. For instance replacing car with cdr should lead to type-level
bugs however none of them are interesting --- they blow up immediately.
Similarly, swapping mutable vectors and hashes with immutable ones is an
interesting mutation in the traditional sense and should lead to a type
error in Typed Racket since its type system distinguishes the two, however
it turns out such mutations  are not interesting in our setting.

* Modes of the rational programmer (section 5).

### 2021-02-10 writing/reviews/pldi-rebuttal-high-level-first.md

> line 90: at first I wondered: what is a case? Maybe say "program
> variant" to foreshadow the idea of generating mutants etc?

It means debugging scenario. We will clarify.

> line 91-ish: "Transient, "Natural", "Erasure" -- be explicit that these
> are names that *you* are introducing

We are reusing these names from Greenman et al. and Vitousek et al. [10, 37],
but will explain more.

@@

> line 380: from this I inferred that "migratory typing" means "gradual
> typing applied at modulewise granularity". Assuming that's correct, it's
> worth saying directly.

We will clarify. Migratory typing is not about granularity, but whether the
type system forces rewrites upon untyped code. A gradual type system for Python
would be migratory. A new gradual-from-the-start language would not be.

> line 424: "fully typed correct programs" -- presumably your method could
> also work with not-yet-fully-typed correct programs, just not ranging
> over the entire lattice in those cases. I was wondering whether that
> might give different/interesting results.

Yes. In fact, we already do this for programs that depend on library code.

### 2021-02-11 writing/reviews/pldi-rebuttal-final.md

PLDI 2021 Paper #66 Author Response
===========================================================================

Thank you to all the reviewers for the helpful feedback.

The first part of our response addresses two common themes in the
reviews; the second part addresses all remaining concerns individually.

Are the empirical results conclusive? Are they surprising?
===========================================================================

Here are the two relevant major conclusions from our experiment:

1. Transient and Natural's blame makes rational programmers more effective than Erasure.

2. The effectiveness of Natural is (merely) comparable to Transient.

As review `D` points out, the second one is surprising: _While Theory
predicts that Natural blames more accurately than Transient, our
experimental results do not support this prediction._

In analogy to physics research, this points to two possible explanations:

- a need for theoretical models with improved predictive power

In all likelihood, the current models are too small to include language features
critical to blame assignment.

- a need for "blame data" from large systems

The presented data is based on systems with up to 14 modules (see table
below).  Natural may prove its worth only when the dependency chains are
deeper than the ones in the existing benchmarks. If Natural works better
for programs with deeper dependency chains, this should guide future research
on mixing approaches.

Is the method novel? 
===========================================================================

Our inspiration is Lazarek et al.'s POPL paper, which shows the possibility
of large-scale design evaluations that do _not_ involve human software devs.

The challenge is how to adapt this work to problems other than
contracts. Given the close relationship between Natural and contracts, it
is "natural" to pursue an adaptation of the Lazarek method to gradual
typing. Adaptation means retaining some ideas and injecting entirely new
ones:

1. Lazarek et al.'s work can use off-the-shelf mutators because they are only
   looking for bugs that contracts can discover in an untyped world.  Our
   work demands an entirely new judgment of which mutators are "good" or
   "bad". This new judgment means cutting one mutator, changing four of them,
   and adding nine. There was no other way to get an _interesting_ and
   diverse body of bugs (see section 4).

2. Lazarek et al. deal with a single contract system. Our experiment compares
   fundamentally different checking and error-reporting schemata: the
   industrial Erasure standard and the academic ones (Natural and Transient).

3. Lazarek et al. examine a single yes-or-no question: whether blame-shifting
   works. Our work asks (a) several yes-or-no questions and (b) asks
   questions of degree; it yields statistically sound answers for the former
   and evidence to shed light on the latter.

4. Lazarek et al.'s method cannot really conclude that blame is the reason
   for the success of the debugging process, because it fails to isolate the
   effect of blame from that of checks or mere luck. It simply validates that
   blame shifting works. Our work focuses on the value of blame and that kind
   of work demands proper controls, which are entirely new.

   Our work relies on two different baselines to understand the value of blame:

   - _the "random" ("lottery") programmer_ ~~ An irrational programmer ignores
     the error report and adds types to one randomly chosen component at a
     time. Hence we use the "random mode" as a baseline to show that following
     the error report is _not_ a question of luck.

   - _the run-time checking of the underlying language_ ~~ All sound gradual
     type systems employ run-time checks to guarantee the consistency of types,
     but they do not have to track blame. If a run-time check fails, they can
     raise an exception that reports a stack (trace).  Many programmers use
     stack traces for debugging, but our work must show that blame assignment
     is better than keeping checks and throwing out blame.

5. One additional point from review `D` stands out, namely, the remark that
   bringing Natural and Transient under the same roof (section 3) is "just
   engineering."

   Neither Reticulated Python nor the model of Transient provide answers
   about what shallow checks correspond to the sophisticated types of Typed
   Racket (or TypeScript), e.g.  polymorphism, row types for mixins,
   occurrence typing) and what information is necessary to record in the
   blame map.  But, we do not think that this paper is the proper place to
   describe this aspect of the work; the large explanation would distract the
   reader from the actual thesis of the paper.

@@

Replies to individual points
==================================================================

Review #66A
===========================================================================
> - Could you comment a bit more on what makes a mutator "good"? (other
>   than experimentally seeing that it is good.) Perhaps, what would be a
>   really bad mutator for evaluating blame assignment?

A bad mutator results in a type-level bug that is trivial to locate. For
instance, a bug that leads to a type-level error during the evaluation of the
component that contains it is trivial to locate. In contrast, a bug that is
non-trivial to locate is one that causes the buggy component to evaluate without
an error and to produce a value of the wrong type. An interesting bug is
one where the value with the wrong type manages also to cross undetected to at
least one other component before a type-level check flags it. Good mutators
should be able to produce these interesting bugs.

@@

> - L991: Figure 6: It took me some time to understand the figure. I
>   wonder if there is a better way to represent the same data?

We have reached the same conclusion and we are considering alternatives
that will make it easier to see how two modes compare to each other, such
as a series of two-sided bar charts.

Review #66B
===========================================================================
> The experiment results show that there are debugging scenarios where
> Transient is more useful than Natural. Does it happen simply because of
> luck? Or, does it imply that Transient is more suitable than Natural for
> a particular kind of bugs for some reason? Also, can there be a new
> blame strategy that outperforms both Transient and Natural all the time?

We were also surprised to discover that Transient outperforms Natural in
some scenarios. After all, theory says otherwise. We haven't identified
any patterns for these scenarios. We conjecture that they point to
limitations of the theory but we can't exclude that some of them aren't
because of some unknown Typed Racket bug.

We do not have any leads for a new blame strategy better than both
Transient and Natural, but at least our method provides a way to examine
ideas in that direction.

@@

> Theoretical gradual type systems usually allow programmers to control
> the precision of type annotations in a fine-grained way with the Dyn
> type rather than distinguishing typed and untyped modules. Would the
> same evaluation method be able to directly appliable to blame in such
> systems?

Yes. Incorporating type Dyn boils down to considering a hierarchy of types
for each component rather than a single type. At the same time, our method
is parametric to the "size" of a component and adapts in a straightforward
manner to a setting where each definition can be considered a component
instead of a whole module. Of course both of these adjustments lead to
significantly larger scenario lattices than the ones we investigate, so
keeping the experiment computationally feasible would require more
aggressive sampling.

> Please explain what impedance mismatches are in the paper. Is it simply
> type mismatches?

"Impedance mismatch" is our name for a run-time disagreement between a
type annotation and an untyped value. The point of the name is to suggest
that either side (type or value) could be at fault. We will adjust the prose.

@@

> The paper states that the effort distribution of the random mode follows
> a normal distribution, as expected. Please explain why a normal
> distribution is a reasonable expectation. ...  This point does not
> affect the contribution of the paper, but it is not clear from the
> paper.

You are right that for a given trail of length $n$ the probability that
the random rational programmer discovers the buggy component on every step
is $1/n$. The expected curve is not flat though, because we randomly
sample from the set of scenarios --- and by extension the set of trails.
That said, as you observe, the distribution of the random programmer's
effort is not relevant to our results, so we will remove the
characterization.

@@

> While the study is mostly well done, it only considers two blame
> assignment strategies.

We would like to point out that the experiment considers six different
modes of the rational programmer spread across three different gradual typing
systems plus the random rational programmer. Figure 7 clearly shows them all.

> For example, it would be interesting to consider the "omniscient
> programmer" to get a bound on the value of blame. But this is not done.

If an "omniscient programmer" is the programmer that always makes the best
choice, i.e. finds the bug at the first try, then this corresponds to a
trivial mode; it has 100% success rate and a constant blame trail length
of 1. Thus, comparing its success to any other strategy amounts to
counting the number of scenarios where the other strategy fails; comparing
effort is similarly trivial. This information can be inferred from the
definitions of section 5, but we can certainly add an explicit mention if
it would be useful.

@@

> And unfortunately the results are inconclusive -- Section 8 says "our
> results call for a deeper understanding of the two models for blame".

The quote from section 8 refers to the theory of blame rather than our
results. Please see the first part of the discussion at the beginning of
the response for more details on the conclusions of our results.

That said, the results of the comparison between the three systems in
section 7 is only one part of the contribution of the paper. The other
part is the method itself (sections 3, 4, and 5).

@@

> Also, the published Transient scheme seems to have a minor bug, which
> might affect the outcomes when fixed (Section 8.2).

Section 8.2 does not discuss a bug in Transient but rather a limitation of
its current design that we discovered exactly because of the experiment.

> To explain one comment above: I was surprised not to see a comparison
> against an "omniscient programmer" to establish a baseline (c.f., the
> SIGPLAN checklist). That is, model a blame assignment scheme in which
> blame could be placed on any component. And always pick the component
> that minimizes the length of the trail. Of course, this would be
> infeasible to implement in practice, but it would place an upper bound
> on the utility of blame.

Our experiment has several baselines. The random mode of the rational
programmer described in section 5.5 serves as the "control" of the whole
experiment. In addition, the exception modes described in sections 5.1 and
5.2 each serve as further "controls" to distinguish the effect of checks
from that of blame within each semantics. Please also see our discussion
of baselines in the second theme at the beginning of our reply.

@@

The Transient semantics provide no interpretation for the blame set. We
picked the first and last because they match the intuitive interpretation
of the blame set as a list. We didn't mean to imply that other options are
unreasonable. We just needed to make a reasonable choice to keep the
experiment feasible.

Review #66D
===========================================================================

> 2. That Transient --- as implemented in Typed Racket --- is much
> slower than Natural. ...
> Given that the experiments are developed and run in the context of the
> large, real-world Racket ecosystem, I wonder whether about the chance of
> incidental implementation choices or bugs affecting this result.

A side note here: Transient Typed Racket is faster than Natural Typed
Racket when we turn off blame. As a sanity check that this is not an
artifact of our implementation, we translated some benchmarks to Python
and confirmed that in Reticulated the performance of Transient with blame
is even worse than we have observed in Typed Racket.

> *Custom Mutators (Section 4)* The paper suggests that the mutators
> required here are significantly different and more subtle, but this does
> not seem to be the case when comparing Figure 3 and Table 3 of the POPL
> 2020 paper.  ...

First of all, we need to clarify that the mutators used by Lazarek et al.
are standard mutators from the the extensive literature on mutation
testing. Naturally, we started with those to develop our mutators, but we
quickly realized that only three of them work for our purposes. In
hindsight, that is not surprising because they are designed to cause
behavioral bugs rather than type-level ones. This realization led us to
think carefully about interesting bugs in our setting and develop a new
set of mutators based on Lazarek et al.'s that meets our needs. For
example, the relational mutator from Lazarek et al. does not lead to
type-level errors that Typed Racket's type system can detect, so we
eliminated it. Similarly, the original arithmetic mutator does not
reliably lead to type errors, so we adapted it to only make operator swaps
that can affect the type of the result. For example, changing a `/` to `*`
will not affect the type of an expression, but changing `*` to `/` may! In
addition to those adjustments, we had to come up with entirely new
mutators in order to get a sufficiently large and diverse set of bugs. For
example, the new `-id` family of mutators are perfectly suited to creating
type-level mistakes, and in fact most of our interesting bugs are
generated by them. These mutators are completely different than swapping
subexpressions; in particular, they are context-sensitive and nonlocal.
Finally, we found mutators that we expected should do well but turned out
to be ineffective at producing interesting bugs. For example, adding a new
public method to a class reliably produces both static and dynamic type
errors, but we found they were not interesting bugs.

Besides all of these details about the mutators, we consider the biggest
insight that allowed us to find the right set of mutators to be the
definition of interesting bugs. Please see the high level comments at the
start for more discussion of this.

> *Debugging Strategies (Section 5)*
>
> These also seem to be overemphasized. The main design choice seems to be
> in how to use Transient blame, which reports multiple components, and
> picking the first and last are pretty intuitive.
>
> I'm also not sure why the exception mode is needed in addition to
> Erasure (I was confused in Section 2 L204-215 and Section 5.1 L732).
> Indeed, Section 5.3 defines the Erasure mode the follow the Natural
> exceptional mode.

The key point of section 5 is the definition of the different blame
strategies in a common precise framework as modes of the rational
programmer. Modes enable the comparison between the different systems even
when sampling is involved, thanks to the common substrate of blame trails.

In addition, modes allow us to introduce in a uniform manner in the
framework baselines that help us isolate confounding factors from the
actual effectiveness of blame. This is exactly the role of the exception
modes for Natural and Transient. Those two perform checks at runtime and
the exception modes help us determine what part of the success of Natural
and Transient is due to the extra checks or due to blame (and also whether
blame masks useful information from exceptions). For example, we see that
Natural without blame but with the same checks outperforms Erasure in ~25%
of the scenarios (figure 6). Blame adds another ~11% on top of that. For
Transient there is a small percentage of scenarios where exceptions do
better than blame but still blame improves over exceptions by ~11%
compared to Erasure.

> *Additional Comments and Typos*
>
> L69: I wish here, and elsewhere, the comparison to [13] had been made
> more explicit. As discussed above, there seem to be more similarities
> than suggested.
> L92: "use of _a_ higher-order contract system"
> L155: "authors extensive"
> L253: "exports _it_ as"
> L280: Missing `untyped-`
> L314: " . "
> L382: "Dyn,"
> Fig 2: Presumably authors were included to indicate the variety of
> sources, but I'm not sure that was necessary.
> L597: "make us_e_ of"
> L664: "described in _Section_ 4"
> L980: "Figure 5.4" ==> "Figure 6"
> L986: "2.8%" ==> "2.18%" ?
> L1017: "33%" ==> "26%" ?
> L1304: "Problem is ...."

@@

Thank you. We will fix these.

Review #66E
===========================================================================

> However, I don't find the results to be as conclusive as the authors
> state. Aside from the Erasure case, whose relevance I found unclear (see
> below), conclusions rest on some small-valued "more useful than"
> percentages. These are pretty hard to interpret, because the
> "usefulness" metrics necessarily build in a lot of simplifying
> assumptions. E.g. in 5.4, the "percentage of scenarios where ... is more
> effective" doesn't account for *extent* of difference, and the
> trail-length "programmer effort" metric in 5.5 is also unlikely to be a
> great proxy. These limitations are understandable, method-wise, but
> overall the results are fairly null... they let us continue believing
> "what we'd expect" but not with any great sense of added confidence.
> That doesn't entirely diminish the work, of course, but I'd say the
> current write-up overplays its hand a little. I appreciate the
> discussion of threats to validity.

Please see the discussion at the beginning of our response about the
interpretation of our results, their significance, and how they challenge
existing perceptions about blame.

We are not sure what "extent" means in this comment. Usefulness asks
whether one system succeeds in a debugging scenario while another system
fails. If both succeed we use the length of the trail to quantify the
difference between the two systems. Trail length may not be the most
optimal proxy for programmer effort, but it also allows us to perform a
large scale automatic experiment about blame without the practical
limitations of user studies. We will add a discussion of this to the
paper.

@@

> The presentation of the results is very much around aggregates and summaries.
> Indeed the whole method is about having run a huge compute job over a large
> number of variants. Perhaps it's paranoia but I'd be interested to see some
> specific examples walked through by hand (in the paper) and some smaller sample
> manually classified (as results). That would add an extra sanity check that the
> metrics do correspond to some meaningful reality.

We are not sure what the question is here. If the question is whether our
implementation of the experiment is faithful to the definitions from
section 5 and the discussion about statistics from section 6, then the
answer is emphatically yes. We have validated manually a number of blame
trails for each mode and moreover our implementation of the experiment
comes with plenty of sanity checks.

If the question is whether our methods acts as a correct classifier for
the detection of bugs a la machine learning, then we would like to stress
that this is not what our work is about. Our work defines the rational
programmer, an ideal programmer, as a deterministic algorithmic procedure
and then uses the results of this procedure on a large number of debugging
scenarios to determine if blame is a critical factor for the success or
failure of the rational programmer.

@@

> About Erasure: if I'm reading this correctly, the key thing here is that
> in Typed Racket, the set of static types is more expressive than the set
> of dynamic types. For example, there is a static notion of non-negative
> integer that is distinct from plain integer. This sort of design isn't
> universal -- in some languages/systems static types closely mirror the
> classification of objects in the language's dynamic semantics. Writing
> these stricter contracts into a program brings its own benefits,
> separate from blame or indeed from static checking. It feels like the
> paper doesn't take enough care to distinguish the two effects: the
> effect of systematically applying more refined contracts over program
> values, and the effect of gradually enabling static checking of those
> contracts (iteratively, guided by blame). The authors do mention this
> around line 204, and the comparisons in Figure 6 between "_ exceptions"
> and "Erasure" seem to be measuring this -- the gain from checking these
> extra annotations, with the more precise and/or more timely checks that
> they imply, relative to the erased case where only the language baseline
> contracts are checked. Indeed that's the point of 'exception'
> experiments. Since the biggest effect sizes on display are these ones --
> between Erasure and anything else -- this seems at best distracting. I'd
> be glad to hear from the authors if I'm misunderstanding anything here.
> In the detailed comments below, I have noted some places in the text
> where it would be useful to remind the reader that this design property
> of Typed Racket is at play.

We think there might be a misunderstanding here about gradual typing:

1. Erasure is the established name of the standard semantics implemented
   by industrial languages such as TypeScript, Flow etc. These languages
   come with a type system just as sophisticated as Typed Racket.
   Programmers add annotations to their program and the type checker does
   its best to type check it in order to discover mistakes and help the
   IDE. Then types are erased and the program is run in the underlying
   dynamically typed language. That language's primitive operations have
   runtime checks that may detect an incompatible argument and raise an
   exception.

2. The two academic semantics introduce extra run time checks (in the form
   of proxies for Natural and inlined checks for Transient, as section 3
   describes). The checks enforce that values the untyped code provides to
   typed code behave according to their expected type. When one of these
   checks fails both Natural and Transient issue blame (though each with
   different mechanisms).

3. In Typed Racket, programmers cannot add contracts. The precision of the
   checks matches the type annotations of the program.

4. The Natural and Transient exception modes perform the same checks as
   Natural and Transient correspondingly. The only difference is that they do
   not produce blame but instead signal an exception with the stack trace of
   the failed check. We use these modes to separate the effectiveness of
   blame from that of the checks. Please also see the second part of the
   discussion at the beginning of our response.
